---
title: "New Residents Patterns in Philadelphia: Part 3"
output: github_document
---

## Data Sources

For this Leading Indicator, we collected data from the IPUMS USA micro data series. We created a data extract from the [IPUMS website](https://usa.ipums.org/usa/) for the following respondent-level characteristics: 

* SEX 
* AGE
* RACE
* HISPAN (Hispanic origin)
* EDUC (Educational attainment)
* INCT (Total personal income)
* MIGPLAC1 (State or county of residence 1 year ago)
* MIGCOUNTY1 (County of residence 1 year ago)
* MIGTYPE1 (Metropolitan status 1 year ago)
* PWSTATE2 (Place of work: state)
* PWCOUNTY (Place of work: county)
* OCC (Occupation)
* TRANWORK (Means of transportation to work)

```{r libraries, include=FALSE}
library(tidyverse)
library(ipumsr)
library(shiny)
library(patchwork)
library(htmltools)
library(DT)
library(ggtext)
library(ggrepel)
library(ggpubr)
library(survey)
```

## Importing Data 

```{r data, include=TRUE}
dat <- read.csv("ipums_nresidents.csv")
```

```{r savingData, include=FALSE}
write_csv(dat, "/Users/elgp/Documents/ELGP_Github/Visuals/2023_Mar_16/ipums_nresidents.csv")
```

## Cleaning Data

```{r analysis, include=FALSE}
# attaching labels to factor variables in IPUMS data 
dat <- dat %>% 
  mutate(MIGPLAC1 = as_factor(MIGPLAC1),
         MIGCOUNTY1 = as_factor(MIGCOUNTY1),
         MIGTYPE1 = as_factor(MIGTYPE1),
         
         PWSTATE2 = as_factor(PWSTATE2),
         PWCOUNTY = as_factor(PWCOUNTY),
         
         RACE = as_factor(RACE),
         RACED = as_factor(RACED),
         SEX = as_factor(SEX),
         EDUC = as_factor(EDUC),
         
         INDNAICS = as_factor(INDNAICS),
         inc_quartile = ntile(INCTOT, 4),
         
         TRANWORK = as_factor(TRANWORK),
         OCC = as_factor(OCC),
         INDNAICS = as_factor(INDNAICS),
         
         HISPAN = as_factor(HISPAN),
         HISPAND = as_factor(HISPAND))
#_________________________________________#
# adding labels to OCCUPATION codes 
dat <- dat %>% 
  mutate(OCC = recode(OCC,
                      `10` = "Chief executives and legislators",
                      `20` = "General and operations managers",
                      `51` = "Marketing managers",
                      `52` = "Sales managers",
                      `102` = "Facilities managers",
                      `110` = "Computer and information systems managers",
                      `120` = "Financial managers",
                      `136` = "Human resources managers",
                      `140` = "Industrial production managers",
                      `150` = "Purchasing managers",
                      `160` = "Transportation, storage, and distribution managers",
                      `220` = "Construction managers",
                      `230` = "Education and childcare administrators",
                      `310` = "Food service managers",
                      `335` = "Entertainment and recreation managers",
                      `350` = "Medical and health services managers",
                      `360` = "Natural sciences managers",
                      `410` = "Property, real estate, and community association managers",
                      `420` = "Social and community service managers",
                      `425` = "Emergency management directors",
                      `440` = "Other managers",
                      `530` = "Purchasing agents, except wholesale, retail, and farm products",
                      `540` = "Claims adjusters, appraisers, examiners, and investigators",
                      `565` = "Compliance officers",
                      `600` = "Cost estimators",
                      `630` = "Human resources workers",
                      `650` = "Training and development specialists",
                      `700` = "Logisticians",
                      `705` = "Project management specialists",
                      `710` = "Management analysts",
                      `726` = "Fundraisers",
                      `735` = "Market research analysts and marketing specialists",
                      `750` = "Business operations specialists, all other",
                      `800` = "Accountants and auditors",
                      `810` = "Property appraisers and assessors",
                      `845` = "Financial and investment analysts",
                      `850` = "Personal financial advisors",
                      `860` = "Insurance underwriters",
                      `930` = "Tax examiners and collectors, and revenue agents",
                      `940` = "Tax preparers",
                      
                      `1006` = "Computer systems analysts",
                      `1010` = "Computer programmers",
                      `1021` = "Software developers",
                      `1032` = "Web and digital interface designers",
                      `1050` = "Computer support specialists",
                      `1108` ="Computer occupations, all other",
                      `1220` ="Operations research analysts",
                      `1240` ="Other mathematical science occupations",
                      `1320` ="Aerospace engineers",
                      `1360` ="Civil engineers",
                      `1420` = "Environmental engineers",
                      `1430` = "Industrial engineers, including health and safety",
                      `1530` = "Other engineers",
                      `1610` = "Biological scientists",
                      `1650` = "Other life scientists",
                      `1720` = "Chemists and materials scientists",
                      `1745` = "Environmental scientists and specialists, including health",
                      `1760` = "Physical scientists, all other",
                      `1860` = "Other social scientists",
                      `1910` = "Biological technicians",
                      `1920` = "Chemical technicians",
                      `1970` = "Other life, physical, and social science technicians",
                      `2001` = "Substance abuse and behavioral disorder counselors",
                      `2002` = "Educational, guidance, and career counselors and advisors",
                      `2004` = "Mental health counselors",
                      `2014` = "Social workers, all other",
                      `2015` = "Probation officers and correctional treatment specialists",
                      `2016` = "Social and human service assistants",
                      `2025` =  "Other community and social service specialists",
                      `2040` = "Clergy",
                      `2050` = "Directors, religious activities and education",
                      `2060` = "Religious workers, all other",
                      `2100` = "Lawyers, and judges, magistrates, and other judicial workers",
                      `2145` = "Paralegals and legal assistants",
                      `2170` = "Title examiners, abstractors, and searchers",
                      `2180` = "Legal support workers, all other",
                      `2205` = "Postsecondary teachers",
                      `2300` = "Preschool and kindergarten teachers",
                      `2310` = "Elementary and middle school teachers",
                      `2320` = "Secondary school teachers",
                      `2330` = "Special education teachers",
                      `2350` = "Tutors",
                      `2360` = "Other teachers and instructors",
                      `2400` = "Archivists, curators, and museum technicians",
                      `2545` = "Teaching assistants", 
                      `2555` = "Other educational instruction and library workers", 
                      `2600` = "Artists and related workers",
                      `2634` = "Graphic designers",
                      `2635` = "Interior designers",
                      `2640` = "Other designers",
                      `2710` = "Producers and directors",
                      `2722` = "Coaches and scouts",
                      `2740` = "Dancers and choreographers",
                      `2752` = "Musicians and singers",
                      `2810` = "News analysts, reporters, and journalists", 
                      `2825` = "Public relations specialists",
                      `2830` = "Editors",
                      `2840` = "Technical writers",
                      `2850` = "Writers and authors",
                      `2861` = "Interpreters and translators",
                      `2910` = "Photographers",
                      `2920` = "Television, video, and film camera operators and editors",
                      `3010` = "Dentists",
                      `3050` = "Pharmacists",
                      `3090` = "Physicians",
                      `3100` = "Surgeons",
                      `3110` = "Physician assistants", 
                      `3140` = "Audiologists",
                      `3150` = "Occupational therapists",
                      `3160` = "Physical therapists",
                      `3230` = "Speech-language pathologists",
                      `3245` = "Other therapists",
                      `3250` = "Veterinarians",
                      `3255` = "Registered nurses",
                      `3256` = "Nurse anesthetists",
                      `3258` = "Nurse practitioners, and nurse midwives",
                      `3300` = "Clinical laboratory technologists and technicians",
                      `3322` = "Diagnostic medical sonographers",
                      `3323` = "Radiologic technologists and technicians",
                      `3421` = "Pharmacy technicians",
                      `3500` = "Licensed practical and licensed vocational nurses", 
                      `3515` = "Medical records specialists",
                      `3550` = "Other healthcare practitioners and technical occupations",
                      `3601` = "Home health aides",
                      `3602` = "Personal care aides",
                      `3603` = "Nursing assistants",
                      `3620` = "Physical therapist assistants and aides",
                      `3630` = "Massage therapists",
                      `3640` = "Dental assistants",
                      `3646` = "Medical transcriptionists",
                      `3649` = "Phlebotomists",
                      `3802` = "Correctional officers and jailers",
                      `3870` = "Police officers",
                      `3930` = "Security guards and gaming surveillance officers",
                      `3960` = "Other protective service workers",
                      `4000` = "Chefs and head cooks",
                      `4010` = "First-line supervisors of food preparation and serving workers",
                      `4020` = "Cooks",
                      `4030` = "Food preparation workers",
                      `4040` = "Bartenders",
                      `4055` = "Fast food and counter workers",
                      `4110` = "Waiters and waitresses",
                      `4130` = "Dining room and cafeteria attendants and bartender helpers",
                      `4140` = "Dishwashers",
                      `4210` = "First-line supervisors of landscaping, lawn service, and groundskeeping workers",
                      `4220` = "Janitors and building cleaners",
                      `4230` = "Maids and housekeeping cleaners",
                      `4255` = "Other grounds maintenance workers",
                      `4350` = "Animal caretakers",
                      `4435` = "Other entertainment attendants and related workers",
                      `4521` = "Manicurists and pedicurists",
                      `4540` = "Tour and travel guides",
                      `4600` = "Childcare workers",
                      `4621` = "Exercise trainers and group fitness instructors",
                      `4622` = "Recreation workers",
                      `4640` = "Residential advisors",
                      `4655` = "Personal care and service workers, all other", 
                      `4700` = "First-line supervisors of retail sales workers",
                      `4710` = "First-line supervisors of non-retail sales workers",
                      `4720` = "Cashiers",
                      `4760` = "Retail salespersons",
                      `4810` = "Insurance sales agents",
                      `4820` = "Securities, commodities, and financial services sales agents", 
                      `4850` = "Sales representatives, wholesale and manufacturing",
                      `4920` = "Real estate brokers and sales agents",
                      `4965` = "Sales and related workers, all other",
                      `5000` = "First-line supervisors of office and administrative support workers",
                      `5100` = "Bill and account collectors",
                      `5120` = "Bookkeeping, accounting, and auditing clerks",
                      `5220` = "Court, municipal, and license clerks",
                      `5230` = "Credit authorizers, checkers, and clerks",
                      `5240` = "Customer service representatives",
                      `5300` = "Hotel, motel, and resort desk clerks", 
                      `5320` = "Library assistants, clerical",
                      `5400` = "Receptionists and information clerks",
                      `5410` = "Reservation and transportation ticket agents and travel clerks",
                      `5510` = "Couriers and messengers",
                      `5730` = "Medical secretaries and administrative assistants", 
                      `5740` = "Secretaries and administrative assistants, except legal, medical, and executive",
                      `5810` = "Data entry keyers",
                      `5820` = "Word processors and typists",
                      `5860` = "Office clerks, general",
                      `5920` = "Statistical assistants",
                      `5940` = "Other office and administrative support workers",
                      `6050` = "Other agricultural workers",
                      `6230` = "Carpenters",
                      `6260` = "Construction laborers",
                      `6355` = "Electricians",
                      `6800` = "Derrick, rotary drill, and service unit operators, and roustabouts, oil, gas, and mining",
                      `7200` = "Automotive service technicians and mechanics",
                      `7315` = "Heating, air conditioning, and refrigeration mechanics and installers",
                      `7330` = "Industrial and refractory machinery mechanics",
                      `7340` = "Maintenance and repair workers, general",
                      `7540` = "Locksmiths and safe repairers",
                      `7610` = "Helpers--installation, maintenance, and repair workers",
                      `7750` = "Other assemblers and fabricators",
                      `7800` = "Bakers",
                      `7810` = "Butchers and other meat, poultry, and fish processing workers",
                      `7840` = "Food batchmakers",
                      `8030` = "Machinists",
                      `8140` = "Welding, soldering, and brazing workers",
                      `8255` = "Printing press operators",
                      `8300` = "Laundry and dry-cleaning workers",
                      `8350` = "Tailors, dressmakers, and sewers",
                      `8610` = "Stationary engineers and boiler operators",
                      `8740` = "Inspectors, testers, sorters, samplers, and weighers",
                      `8760` = "Dental and ophthalmic laboratory technicians and medical appliance technicians",
                      `8800` = "Packaging and filling machine operators and tenders",
                      `9050` = "Flight attendants",
                      `9130` = "Driver/sales workers and truck drivers",
                      `9142` = "Taxi drivers",
                      `9300` = "Sailors and marine oilers, and ship engineers",
                      `9310` = "Ship and boat captains and operators", 
                      `9415` = "Passenger attendants",
                      `9610` = "Cleaners of vehicles and equipment", 
                      `9620` = "Laborers and freight, stock, and material movers, hand",
                      `9640` = "Packers and packagers, hand",
                      `9645` = "Stockers and order fillers",
                      `9720` = "Refuse and recyclable material collectors",
                      `9830` = "Military, rank not specified"
                      ))
```

### Attaching survey weights

```{r weights, include=TRUE}
#_________________________________________#
# separating data into pre and post pandemic 
pre_COVID <- dat %>% 
  filter(YEAR == 2018 | YEAR == 2019)
post_COVID <- dat %>% 
  filter(YEAR == 2020 | YEAR == 2021)
#_________________________________________#
# attaching survey weights 
dtaDesign_pre <- svydesign(id      = ~CLUSTER,
                           strata  = ~STRATA,
                           weights = ~PERWT,
                           nest    = TRUE,
                           data    = pre_COVID)
#_________________________________________#
dtaDesign_post <- svydesign(id     = ~CLUSTER,
                           strata  = ~STRATA,
                           weights = ~PERWT,
                           nest    = TRUE,
                           data    = post_COVID)
```

## Cross Tabulations 

### 1. Top 5 Occupations 

```{r occupations_tab, include=TRUE}
# 1. Occupation => describe the top 10 jobs held by migrant workers 
table <- svytable(~OCC+YEAR, design = dtaDesign_pre)
df_pre <- as.data.frame.matrix(table) 
df_pre <- rownames_to_column(df_pre, var = "Variable") %>% as_tibble()
#_________________________________________#
table <- svytable(~OCC+YEAR, design = dtaDesign_post)
df_post <- data.frame(rbind(table))
df_post <- rownames_to_column(df_post, var = "Variable") %>% as_tibble()
#_________________________________________#
df_occ <- df_pre %>% 
  left_join(df_post, by = c("Variable")) %>% 
  filter(Variable != "")
# calculating top 10 OCCUPATIONS that new residents are working in 
df_occ <- df_occ %>% 
  gather(YEAR, count, `2018`:X2021) %>% 
  group_by(YEAR) %>% 
  arrange(desc(count)) %>%
  group_by(YEAR) %>% 
  mutate(prop_count = count/sum(count)) %>% 
  slice(1:10) %>% 
  mutate(Variable = ifelse(Variable == "Secretaries and administrative assistants, except legal, medical, and executive", "Secretaries and administerative assistants", 
                           ifelse(Variable == "Lawyers, and judges, magistrates, and other judicial workers", "Lawyers, judges and judicial workers", Variable)))
```

### 2. Are New Residents Mostly Working in Philadelphia?

``` {r workplace, include = TRUE}
###########################################
# 6. Place of work => can point out whether numbers of migrants working outside PA 
# has increased since the pandemic or not
# separating data into pre and post pandemic 
#:::::::::::::::::::::::::::::::::::::::::
table <- svytable(~PWCOUNTY+YEAR, design = dtaDesign_pre)
df_pre <- as.data.frame.matrix(table) 
df_pre <- rownames_to_column(df_pre, var = "Variable") %>% as_tibble()
df_pre <- df_pre %>% 
  group_by(Variable) %>% 
  mutate(`Pre-COVID` = mean(`2018`:`2019`))
#_________________________________________#
table <- svytable(~PWCOUNTY+YEAR, design = dtaDesign_post)
df_post <- data.frame(rbind(table))
df_post <- rownames_to_column(df_post, var = "Variable") %>% as_tibble() 
df_post <- df_post %>% 
  group_by(Variable) %>% 
  mutate(`Post-COVID` = mean(X2020:X2021))
#_________________________________________#
df_pwork <- df_pre %>% 
  left_join(df_post, by = c("Variable"))
#_________________________________________#
# creating dummy of philly 
df_pwork <- df_pwork %>% 
  select(Variable, `2018`, `2019`, X2020, X2021) %>% 
  gather(Year, count, `2018`:X2021) %>% 
  mutate(Philly = ifelse(Variable == 101, "Philadelphia", "Not Philadelphia")) %>% 
  group_by(Year, Philly) %>% 
  mutate(sum_pwork = sum(count)) %>%
  group_by(Year) %>% 
  mutate(prop_pwork = sum_pwork/sum(count))
# creating lineplots of changing proportions in migrant workers working in philadelphia vs outside
df_pwork <- df_pwork %>% 
  mutate(Year = recode(Year,
                       "2018" = 2018,
                       "2019" = 2019,
                       "X2020" = 2020,
                       "X2021" = 2021)) %>% 
  distinct(prop_pwork, .keep_all = TRUE) %>% 
  group_by(Philly) %>% 
  mutate(index_prop = prop_pwork/prop_pwork[which(Year == 2018)])
# adding labels to maximum year in dataset 
df_pwork <- df_pwork %>% 
  mutate(label = ifelse(Philly == "Not Philadelphia" & Year == 2018, "Working Outside Philadelphia",
                        ifelse(Philly == "Philadelphia" & Year == 2018, "Working Inside Philadelphia", NA)))
#:::::::::::::::::::#
#       2018
#:::::::::::::::::::#
# RECORDING PWCOUNTIES THAT CORRESPOND TO PWSTATES 
# COUNTY FIPS CODE: 91 => Montgomery County, PA
# COUNTY FIPS CODE: 7 => Camden County, NJ
# COUNTY FIPS CODE: 29 => Chester County, PA
# COUNTY FIPS CODE: 45 => Delaware County, PA
# COUNTY FIPS CODE: 17 => Bucks County, PA
#:::::::::::::::::::#
#       2020
#:::::::::::::::::::#
# RECORDING PWCOUNTIES THAT CORRESPOND TO PWSTATES 
# COUNTY FIPS CODE: 91 => Montgomery County, PA
# COUNTY FIPS CODE: 3 => New Castle, DE
# COUNTY FIPS CODE: 7 => Camden County, NJ
# COUNTY FIPS CODE: 15 => Gloucester County, NJ
# COUNTY FIPS CODE: 45 => Delaware County, PA
#:::::::::::::::::::#
#       2021
#:::::::::::::::::::#
# RECORDING PWCOUNTIES THAT CORRESPOND TO PWSTATES 
# COUNTY FIPS CODE: 61 => New York County, PA
# COUNTY FIPS CODE: 7 => Camden County, NJ
# COUNTY FIPS CODE: 91 => Montgomery County, PA
# COUNTY FIPS CODE: 17 => Bucks County, PA
#:::::::::::::::::::::::::::::::::::::::::
###########################################
```

### 3. Is Remote Work Higher among New Residents After COVID-19

```{r remotework, include = TRUE}
# 7a. Remote work => is remote work higher in migrants who came to Philly after the pandemic
table <- svytable(~TRANWORK+YEAR, design = dtaDesign_pre)
df_pre <- as.data.frame.matrix(table) 
df_pre <- rownames_to_column(df_pre, var = "Variable") %>% as_tibble() 
#_________________________________________#
table <- svytable(~TRANWORK+YEAR, design = dtaDesign_post)
df_post <- data.frame(rbind(table))
df_post <- rownames_to_column(df_post, var = "Variable") %>% as_tibble()
#_________________________________________#
df_remote <- df_pre %>% 
  left_join(df_post, by = c("Variable"))
#_________________________________________#
# calculating proportion of remote work among new residents
df_remote <- df_remote %>% 
  gather(Year, count, `2018`:X2021) %>% 
  group_by(Year) %>% 
  mutate(remote_prop = count[which(Variable == "80")]/sum(count) # remote work code = 80
  )
```

#### 4. What are the top jobs among new residents who work remotely?

``` {r remotejobs, include = TRUE}
# 7b. Remote work jobs => what are the top 10 jobs that remote working migrants have?
#_________________________________________#
table <- svytable(~TRANWORK+OCC, design = dtaDesign_post)
df_post <- as.data.frame(table)
df_post <- df_post %>% rename(FREQ_postCOVID = Freq)
#_________________________________________#
df_remoteOCC <- df_post %>% 
  filter(TRANWORK == "80") # filtering for remote work only
#_________________________________________#
df_remoteOCC <- df_remoteOCC %>%
  mutate(OCC = as.character(OCC)) %>% 
  mutate(OCC = ifelse(OCC == "Other community and social service specialists", "Other social sercvice specialists",
                      ifelse(OCC == "Real estate brokers and sales agents", "Real estate brokers",
                             ifelse(OCC == "Transportation, storage, and distribution managers", "Transportation managers", OCC))))
#_________________________________________#
# calculating proportions of remote work occupations 
df_remoteOCC <- df_remoteOCC %>% 
  mutate(prop_postCOVID = FREQ_postCOVID/sum(FREQ_postCOVID)
         ) %>% 
  arrange(desc(prop_postCOVID)) %>% # choosing top 10 remote work occupations by proportion
  slice(1:10)
```

